---
title: "Report Card Scoring"
author: "Tampa Bay Estuary Program"
vignette: >
  %\VignetteIndexEntry{scoring}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
    echo: true
    warning: false
    message: false
---

```{r}
#| label: setup
#| include: false
library(tbepreport)
```

This article describes the full workflow for creating the bay segment annual report card. This includes descriptions of computing each
individual indicator, converting raw indicator values to a common 0-1
outcome scale, combining indicators into category scores and an overall
bay segment score, and visualizing the results. The four categories -
Water Quality, Sediment, Fish and Wildlife, and Habitat - are each covered
below, followed by sections on scoring, combining, and plotting.

## Water Quality

### Targets

Combines chlorophyll and light attenuation attainment into a single
continuous outcome: the two sub-scores are summed, so a lower combined
sub-score (better attainment on both fronts) gives a higher outcome.  These
outcomes come directly from the management targets for each bay segment, 
as described in the tbeptools [water quality vignette](https://tbep-tech.github.io/tbeptools/articles/intro.html){target="_blank"}.

```{r}
wqattain <- anlz_wq_attain(epcdata)
```

### Chlorophyll and Light Thresholds

Compares annual mean chlorophyll and light attenuation values directly
against bay-segment-specific thresholds, independent of the combined
attainment score above.  Each gives its own threshold outcome as a smooth
logistic transition by default or a hard binary 0/1 with `smooth = FALSE` 
(see [Scoring](#scoring)).

```{r}
wqthresh <- anlz_wq_thresh(epcdata)
```

### Nutrient and Hydrologic Loading

Compares total nitrogen load and a hydrologically-normalized loading
against fixed bay-segment thresholds, each as its own threshold outcome -
a smooth logistic transition by default, or a hard binary 0/1 with
`smooth = FALSE` - see [Scoring](#scoring).

```{r}
totanndat <- util_rdataload("https://github.com/tbep-tech/load-estimates/raw/main/data/totanndat.RData")
wqload <- anlz_wq_load(totanndat)
```

### Tidal Creeks

Assigns each tidal creek to a bay segment, converts its condition category
(Prioritize/Investigate/Caution/Monitor) to an outcome, then averages by
bay segment/year weighted by creek length so longer creek segments
contribute proportionally more.

```{r}
wqtidalcreeks <- anlz_wq_tidalcreeks(tidalcreeks, iwrraw, yrs = 1975:2024)
```

### Fecal Indicator Bacteria

Scores each bay segment/year from enterococcus monitoring data using
`exceed_rate`, a continuous, one-sided 90% upper confidence estimate of
the true exceedance rate (0-1, lower is better), rescaled to an outcome
as `1 - exceed_rate`. This is a continuous analog of the A-E letter grade
the same underlying calculation also produces (`cat`), which is itself a
discretized version of `exceed_rate`.

```{r}
wqfib <- anlz_wq_fib(enterodata)
```

## Sediment

### Contaminants

Grades each bay segment/year A-F from its average sediment contamination
score (PEL/TEL), then converts the letter grade to an outcome.

```{r}
peltel <- anlz_sed_peltel(sedimentdata, yrs = 1993:2024)
```

### Benthic Index

Grades each bay segment/year Poor/Fair/Good from the Tampa Bay Benthic
Index (TBBI), then converts the grade to an outcome.

```{r}
tbbi <- anlz_sed_tbbi(benthicdata)
```

## Fish and Wildlife

### Nekton Index

Scores each bay segment/year 0-100 with the Tampa Bay Nekton Index (TBNI),
then rescales to an outcome using TBNI's own grade breakpoints (On Alert
below 32, Caution from 32 to 46, Stay the Course above 46): scores below 32
give an outcome of 0, scores above 46 give an outcome of 1, and scores in
between are linearly rescaled.

```{r}
tbni <- anlz_fw_tbni(fimdata)
```

### Non-natives

Ports the non-native species "report card" metrics from the
[tbep-invasives](https://github.com/tbep-tech) Python pipeline
(`src/tbep_invasives/steps/report_cards.py`) to R. Returns non-native species
abundance (observations per unit area per year) and richness (unique
species per unit area per year) by bay segment, converted to percentiles
and then reversed to an outcome (higher abundance/richness is worse).

```{r}
nonnative_obs <- anlz_fw_nonnative_obs()
nonnative_abundance <- anlz_fw_nonnative_abundance(nonnative_obs)
nonnative_richness  <- anlz_fw_nonnative_richness(nonnative_obs)
```

## Habitat

### Seagrass Transects

Estimates frequency of seagrass occurrence (0-100) at each bay
segment/year from transect monitoring data, then linearly rescales to an
outcome.

```{r}
trnsct <- anlz_hab_seagrass_transect(transect)
```

### Seagrass Coverage

Compares mapped seagrass acreage against a fixed per-segment target (a
threshold outcome - a smooth logistic transition by default, or a hard
binary 0/1 with `smooth = FALSE` - see [Scoring](#scoring)), where the
targets are the baywide 40,000-acre seagrass coverage target apportioned
across segments by each segment's share of total bay area. Coverage maps
are flown only every couple of years, so non-survey years carry forward
the most recent actual estimate and outcome.

```{r}
cov <- anlz_hab_seagrass_coverage(sgsegest)
```

## Scoring

Every indicator above ends with the same step: a raw measurement is
converted to a 0-1 outcome, where 1 is always the best possible condition.
That conversion happens in one place, `util_outcome()`, so how an
indicator is scored can change without touching the `anlz_*` function that
uses it. `util_outcome()` supports three types:

* **Continuous** - a raw value is linearly rescaled to 0-1 over a specified
  range, clamped so values outside that range are pinned to 0 or 1 rather
  than extrapolated.  This can occur for the full range of raw values
  (e.g., seagrass transect frequency of occurrence over
  its full 0-100 range) or within a subset range of the raw values (e.g., the Nekton Index's 0-100 score rescaled over just its 32-46 breakpoint window).
  FIB's `exceed_rate` is also scored this way, reversed since lower is
  better (`from = c(0, 1)`, `reverse = TRUE`).
* **Category** - a discrete grade or condition category is mapped to a
  fixed outcome (e.g. sediment PEL/TEL grades A-F, tidal creek condition
  categories, TBBI Poor/Fair/Good).
* **Threshold** - a raw value is compared against a cutoff (e.g. nutrient
  loading against a bay-segment target, chlorophyll/light attenuation
  against their thresholds, seagrass coverage against an acreage target).
  By default this gives a smooth logistic outcome: exactly 0.5 at the
  threshold, moving quickly toward 0 or 1 (depending on direction) as a
  value moves away from it, so values close to the threshold are
  penalized less harshly than a hard cutoff would. The steepness of that
  transition is set by `pct`, a fraction of the threshold itself (default
  10%) rather than a fixed absolute value, so it stays meaningful across
  indicators/segments with very different thresholds. Setting
  `smooth = FALSE` instead gives a hard binary 0 or 1.

```{r}
# continuous: linearly rescale a raw value between known bounds
util_outcome(75, type = 'continuous', from = c(0, 100))

# category: map a discrete grade to a fixed outcome
util_outcome('B', type = 'category', levels = c(A = 1, B = 0.75, C = 0.5, D = 0.25, F = 0))

# threshold: smooth by default, a logistic transition centered on the cutoff
util_outcome(8, type = 'threshold', thresh = 10, op = '<')

# threshold, smooth = FALSE: a hard binary cutoff instead
util_outcome(8, type = 'threshold', thresh = 10, op = '<', smooth = FALSE)
```

### Continuous Example

The plot below compares how the Nekton Index's 0-100 score is rescaled to
an outcome under the current default (`from = c(32, 46)`, TBNI's own
grade breakpoints - clamped so anything below 32 is 0 and above 46 is 1)
versus how it was previously scored (`from = c(0, 100)`, a plain linear
rescale over the full range, with no clamping needed since no score falls
outside 0-100). Each bay segment/year's actual TBNI score (from `tbni`,
computed earlier) is overlaid on the current curve - most observed scores
fall inside or above the 32-46 breakpoint window, where the two methods
diverge most.

```{r}
#| echo: false
xvals <- seq(0, 100, by = 0.5)
new_outcome <- util_outcome(xvals, type = 'continuous', from = c(32, 46))
old_outcome <- util_outcome(xvals, type = 'continuous', from = c(0, 100))

plot(xvals, new_outcome, type = 'l', lwd = 2, col = 'steelblue', ylim = c(0, 1),
     xlab = 'TBNI score (0-100)', ylab = 'Outcome',
     main = 'TBNI scoring: 32-46 breakpoint vs. plain 0-100 rescale')
lines(xvals, old_outcome, lwd = 2, lty = 2, col = 'grey40')
abline(v = c(32, 46), lty = 3, col = 'grey70')
points(tbni$Segment_TBNI, tbni$outcome, pch = 16, col = 'darkorange')
legend('bottomright',
       legend = c('32-46 breakpoint (current default)', 'plain 0-100 rescale', 'hypothetical segment/year scores'),
       col = c('steelblue', 'grey40', 'darkorange'), lwd = c(2, 2, NA), lty = c(1, 2, NA), pch = c(NA, NA, 16), bty = 'n')
```

### Smooth vs. Hard Threshold Example

The plot below compares the hard cutoff to the smooth logistic outcome at
a few `pct` values, for an arbitrary threshold of 10: all three smooth
curves cross 0.5 exactly at the threshold, and a larger `pct` widens the
transition around it.

```{r}
#| echo: false
xvals <- seq(0, 20, by = 0.1)
thresh <- 10

plot(xvals, as.numeric(xvals < thresh), type = 's', lwd = 2, col = 'grey40',
     xlab = 'Raw value', ylab = 'Outcome', main = 'Hard cutoff vs. smooth threshold (thresh = 10)')
lines(xvals, util_outcome(xvals, type = 'threshold', thresh = thresh, smooth = TRUE, pct = 0.05), col = 'steelblue', lwd = 2)
lines(xvals, util_outcome(xvals, type = 'threshold', thresh = thresh, smooth = TRUE, pct = 0.1), col = 'darkorange', lwd = 2)
lines(xvals, util_outcome(xvals, type = 'threshold', thresh = thresh, smooth = TRUE, pct = 0.2), col = 'firebrick', lwd = 2)
abline(v = thresh, lty = 2, col = 'grey70')
legend('right', legend = c('hard cutoff', 'pct = 0.05', 'pct = 0.1', 'pct = 0.2'),
       col = c('grey40', 'steelblue', 'darkorange', 'firebrick'), lwd = 2, bty = 'n')
```

Applying this to Old Tampa Bay's chlorophyll threshold
(9.3 ug/L) demonstrates how many observed annual means actually fall
within the default 10% transition band: about two-thirds of OTB's 52
years of data land between outcomes of 0.1 and 0.9, since a value tracked
against a management target tends to hover close to it. For an indicator
like this, the default smooth scoring meaningfully changes scoring for a
majority of years, not just a few borderline ones (compared to
`smooth = FALSE`):

```{r}
#| echo: false
otb_chla <- tbeptools::anlz_avedat(epcdata)$ann |>
  dplyr::filter(bay_segment == 'OTB', var == 'mean_chla')
otb_thresh <- tbeptools::targets$chla_thresh[tbeptools::targets$bay_segment == 'OTB']

xvals <- seq(0, 20, by = 0.1)
plot(xvals, util_outcome(xvals, type = 'threshold', thresh = otb_thresh, op = '<', smooth = TRUE),
     type = 'l', lwd = 2, col = 'steelblue',
     xlab = 'Mean chlorophyll (ug/L)', ylab = 'Outcome',
     main = 'OTB chlorophyll: smooth threshold vs. observed values')
abline(v = otb_thresh, lty = 2, col = 'grey70')
points(otb_chla$val, util_outcome(otb_chla$val, type = 'threshold', thresh = otb_thresh, op = '<', smooth = TRUE),
       pch = 16, col = 'darkorange')
legend('topright', legend = c('smooth threshold curve', 'observed annual means'),
       col = c('steelblue', 'darkorange'), lwd = c(2, NA), pch = c(NA, 16), bty = 'n')
```

For comparison, here is the same OTB chlorophyll data scored with `smooth = FALSE` - a hard binary cutoff instead of the logistic transition above. Every point lands on exactly 0 or 1, regardless of how close its value is to the 9.3 ug/L threshold:

```{r}
#| echo: false
plot(xvals, util_outcome(xvals, type = 'threshold', thresh = otb_thresh, op = '<', smooth = FALSE),
     type = 's', lwd = 2, col = 'grey40',
     xlab = 'Mean chlorophyll (ug/L)', ylab = 'Outcome',
     main = 'OTB chlorophyll: hard cutoff vs. observed values')
abline(v = otb_thresh, lty = 2, col = 'grey70')
points(otb_chla$val, util_outcome(otb_chla$val, type = 'threshold', thresh = otb_thresh, op = '<', smooth = FALSE),
       pch = 16, col = 'darkorange')
legend('right', legend = c('hard cutoff', 'observed annual means'),
       col = c('grey40', 'darkorange'), lwd = c(2, NA), pch = c(NA, 16), bty = 'n')
```

## Combining Scores

Once every indicator has a 0-1 outcome, three functions combine them into
increasingly aggregated scores. The diagram below shows the full
hierarchy for a single bay segment/year: each category's indicators
(named as they're passed into `anlz_indicators()`/`anlz_category()` below)
funnel into that category's score, and the four category scores funnel
into the single overall score.

```{mermaid}
%%| fig-width: 6
flowchart LR
    A1["wq_attain"] --> WQ["Water Quality"]
    A2["thresh"] --> WQ
    A3["load"] --> WQ
    A4["tidal_creeks"] --> WQ
    A5["fib"] --> WQ

    B1["sed_peltel"] --> SED["Sediment"]
    B2["sed_tbbi"] --> SED

    C1["tbni"] --> FW["Fish and Wildlife"]
    C2["nonnative_abundance"] --> FW
    C3["nonnative_richness"] --> FW

    D1["seagrass_transect"] --> HAB["Habitat"]
    D2["seagrass_coverage"] --> HAB

    WQ --> SCORE["Bay Segment / Year Score"]
    SED --> SCORE
    FW --> SCORE
    HAB --> SCORE

    classDef indicator fill:#eef2f7,stroke:#8aa1b1,color:#1c1c1c;
    classDef category fill:#ffe8b3,stroke:#c98a12,color:#1c1c1c;
    classDef score fill:#c8e6c9,stroke:#2e7d32,color:#1c1c1c;

    class A1,A2,A3,A4,A5,B1,B2,C1,C2,C3,D1,D2 indicator;
    class WQ,SED,FW,HAB category;
    class SCORE score;
```

* `anlz_indicators()` stacks any number of named indicator data.frames into
  one long table (`bay_segment`, `yr`, `indicator`, `outcome`), without
  filling in gaps where an indicator has no data for a given segment/year.
* `anlz_category()` calls `anlz_indicators()` internally, then averages
  `outcome` within each bay segment/year to get a category score - the
  result also keeps one wide column per indicator, so the indicators behind
  a category score stay visible alongside it.
* `anlz_score()` is a thin wrapper around `anlz_category()`: the four
  category scores are themselves just another set of "indicators" to
  average, giving the overall bay segment score.

```{r}
wqoverall <- anlz_category(
  wq_attain = wqattain, 
  thresh = wqthresh, 
  load = wqload, 
  tidal_creeks = wqtidalcreeks, 
  fib = wqfib
)
```

```{r}
sedoverall <- anlz_category(
  sed_peltel = peltel, 
  sed_tbbi = tbbi
)
```

```{r}
fwoverall <- anlz_category(
  tbni                 = tbni,
  nonnative_abundance  = nonnative_abundance,
  nonnative_richness   = nonnative_richness
)
```

```{r}
haboverall <- anlz_category(
  seagrass_transect = trnsct,
  seagrass_coverage = cov
)
```

```{r}
score <- anlz_score(wqoverall, sedoverall, fwoverall, haboverall)
```

## Visualizing Scores

`plot_category()` and `plot_score()` both plot a two- or three-ring
sunburst - a colored center hole for the overall score, surrounded by a
ring per component, each colored on the same continuous red/yellow/green
outcome scale. `plot_category()` takes the same named indicator
data.frames as `anlz_category()` (and calls it internally) to plot one
category's indicators around its score, for a chosen bay segment and year:

```{r}
plot_category(
  wq_attain    = wqattain,
  thresh       = wqthresh,
  load         = wqload,
  tidal_creeks = wqtidalcreeks,
  fib          = wqfib,
  bay_segment = 'OTB', yr = 2024
)
```

```{r}
plot_category(sed_peltel = peltel, sed_tbbi = tbbi, bay_segment = 'OTB', yr = 2024)
```

```{r}
plot_category(
  tbni                = tbni,
  nonnative_abundance = nonnative_abundance,
  nonnative_richness  = nonnative_richness,
  bay_segment = 'OTB', yr = 2024
)
```

```{r}
plot_category(
  seagrass_transect = trnsct,
  seagrass_coverage = cov,
  bay_segment = 'OTB', yr = 2024
)
```

`plot_score()` takes the same four category data.frames as `anlz_score()`
(and calls it internally) to plot the full hierarchy - indicators, then
categories, then the overall score - for a chosen bay segment and year:

```{r}
plot_score(wqoverall, sedoverall, fwoverall, haboverall, bay_segment = 'OTB', yr = 2024)
```

Unlike the sunburst plots above, which show one bay segment/year at a time,
`plot_trend()` shows every year at once for a chosen bay segment: five
stacked facets, an "Overall" facet with the bay segment score and one line
per category, followed by one facet per category with that category's own
score and one line per indicator. In every facet the score itself is a
bold dark line, and its components are thinner colored lines labeled
directly at their right-most point (rather than a single shared legend,
which would otherwise have to list every category and indicator name at
once):

```{r}
#| fig-height: 12
plot_trend(wqoverall, sedoverall, fwoverall, haboverall, bay_segment = 'OTB', 
  yr_range = c(2000, 2024))
```

## Notes

* Indices that use categories from a continuous scale, revert to continuous data to get outcome
* Lots of redundancy in the water quality indicators
* How to incorporate land use change by bay segment? Is this even appropriate?
