---
title: "Trends and Sensitivity"
author: "Tampa Bay Estuary Program"
vignette: >
  %\VignetteIndexEntry{trends}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
    echo: true
    warning: false
    message: false
---

```{r}
#| label: setup
#| include: false
library(tbepreport)
library(ggplot2)
```

This article has two parts. The first compares report card results across
bay segments, both as trends from 2000-2024 and as a snapshot of 2024
conditions. The second shows how scoring decisions (highlighted in
the [Report Card Scoring](scoring.html) article) can change the results,
using the Nekton Index's rescaling window, threshold smoothing, and
category weights as examples.

## Setup

Indicators and category scores are built the same way as in
[Report Card Scoring](scoring.html), for the four default bay segments
(`OTB`, `HB`, `MTB`, `LTB`).

```{r}
bay_segments <- c('OTB', 'HB', 'MTB', 'LTB')

wqattain <- anlz_wq_attain(epcdata)
wqthresh <- anlz_wq_thresh(epcdata)
totanndat <- util_rdataload("https://github.com/tbep-tech/load-estimates/raw/main/data/totanndat.RData")
wqload <- anlz_wq_load(totanndat)
wqtidalcreeks <- anlz_wq_tidalcreeks(tidalcreeks, iwrraw, yrs = 1990:2024)
wqfib <- anlz_wq_fib(enterodata)

peltel <- anlz_sed_peltel(sedimentdata, yrs = 1993:2024)
tbbi <- anlz_sed_tbbi(benthicdata)

tbni <- anlz_fw_tbni(fimdata)
nonnative_obs <- anlz_fw_nonnative_obs()
nonnative_abundance <- anlz_fw_nonnative_abundance(nonnative_obs)
nonnative_richness  <- anlz_fw_nonnative_richness(nonnative_obs)

trnsct <- anlz_hab_seagrass_transect(transect)
cov <- anlz_hab_seagrass_coverage(sgsegest)

wqoverall <- anlz_category(
  wq_attain    = wqattain,
  thresh       = wqthresh,
  load         = wqload,
  tidal_creeks = wqtidalcreeks,
  fib          = wqfib
)
sedoverall <- anlz_category(
  sed_peltel = peltel,
  sed_tbbi   = tbbi
)
fwoverall <- anlz_category(
  tbni                = tbni,
  nonnative_abundance = nonnative_abundance,
  nonnative_richness  = nonnative_richness
)
haboverall <- anlz_category(
  seagrass_transect = trnsct,
  seagrass_coverage = cov
)
```

## Part 1: Comparing Bay Segments

### Trends by Bay Segment

Each panel below shows one bay segment's overall score from 2000-2024
(the "Overall" facet from `plot_trend()`). Labels are turned off since
they would collide at this smaller panel size.

```{r}
#| layout-ncol: 2
for (seg in bay_segments) {
  p <- plot_trend(
    wqoverall, sedoverall, fwoverall, haboverall, bay_segment = seg,
    yr_range = c(2000, 2024), facets = 'Overall', labels = FALSE
  )
  print(p)
}
```

### 2024 Snapshot by Bay Segment

Each sunburst below shows the same 2024 outcome scale across all four bay
segments.  Color ranges are the same across plots for visual comparison of 
score ranges.

::: {layout-ncol="2"}
```{r}
plot_score(wqoverall, sedoverall, fwoverall, haboverall, bay_segment = 'OTB', yr = 2024)
```

```{r}
plot_score(wqoverall, sedoverall, fwoverall, haboverall, bay_segment = 'HB', yr = 2024)
```

```{r}
plot_score(wqoverall, sedoverall, fwoverall, haboverall, bay_segment = 'MTB', yr = 2024)
```

```{r}
plot_score(wqoverall, sedoverall, fwoverall, haboverall, bay_segment = 'LTB', yr = 2024)
```
:::

## Part 2: How Scoring Decisions Affect Results

### Nekton Index: Breakpoint Window vs. Full Range

`anlz_fw_tbni()` defaults to rescaling the Nekton Index's 0-100 score over
its own 32-46 grade breakpoints (`from = c(32, 46)`). Passing
`from = c(0, 100)` instead spreads sensitivity evenly across the full
range rather than concentrating it near the breakpoints.

```{r}
#| layout-ncol: 2
tbni_full <- anlz_fw_tbni(fimdata, from = c(0, 100))
fwoverall_full <- anlz_category(
  tbni                = tbni_full,
  nonnative_abundance = nonnative_abundance,
  nonnative_richness  = nonnative_richness
)

plot_trend(
  wqoverall, sedoverall, fwoverall, haboverall, bay_segment = 'OTB',
  facets = 'fw', labels = FALSE, yr_range = c(2000, 2024)
) +
  labs(subtitle = '32-46 breakpoint (default)')

plot_trend(
  wqoverall, sedoverall, fwoverall_full, haboverall, bay_segment = 'OTB',
  facets = 'fw', labels = FALSE, yr_range = c(2000, 2024)
) +
  labs(subtitle = 'Full 0-100 range')
```

### Threshold Scoring: Smooth vs. Hard Cutoff

`anlz_wq_thresh()` defaults to a smooth logistic transition at each bay
segment's chlorophyll/light attenuation threshold (`smooth = 'logistic'`).
`smooth = 'none'` instead uses a hard 0/1 cutoff, as described in
[Report Card Scoring](scoring.html#smooth-vs.-hard-threshold-example).

```{r}
#| layout-ncol: 2
wqthresh_hard <- anlz_wq_thresh(epcdata, smooth = 'none')
wqoverall_hard <- anlz_category(
  wq_attain    = wqattain,
  thresh       = wqthresh_hard,
  load         = wqload,
  tidal_creeks = wqtidalcreeks,
  fib          = wqfib
)

plot_trend(
  wqoverall, sedoverall, fwoverall, haboverall, bay_segment = 'OTB',
  facets = 'wq', labels = FALSE, yr_range = c(2000, 2024)
) +
  labs(subtitle = 'Smooth logistic (default)')

plot_trend(
  wqoverall_hard, sedoverall, fwoverall, haboverall, bay_segment = 'OTB',
  facets = 'wq', labels = FALSE, yr_range = c(2000, 2024)
) +
  labs(subtitle = 'Hard cutoff (smooth = "none")')
```

### Category Weights

`anlz_score()`'s `wt` argument weights the four category scores when
combining them into the overall score (default `NULL`, equal weight).
Below, Water Quality is upweighted twice relative to the other three
categories.

```{r}
#| layout-ncol: 2
plot_trend(
  wqoverall, sedoverall, fwoverall, haboverall, bay_segment = 'OTB',
  facets = 'Overall', labels = FALSE, yr_range = c(2000, 2024)
) +
  labs(subtitle = 'Equal category weights (default)')

plot_trend(
  wqoverall, sedoverall, fwoverall, haboverall, bay_segment = 'OTB',
  facets = 'Overall', labels = FALSE, wt = c(wq = 2, sed = 1, fw = 1, hab = 1), 
  yr_range = c(2000, 2024)
) +
  labs(subtitle = 'Water Quality weighted 2x')
```
